Ethical considerations when planning, implementing and releasing health economic model software: a new proposal

### 2024-03-20 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
ready for preprint

---
title: "MS_Aims"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
 <!-- Important rewordings --> 
In this paper, we <!--  MOVED_EDITED_START -->draw on our prior experience with developing and using CHEMs and consider literature on modelling practice to propose<!--  MOVED_EDITED_END -->:  (i) a set of ethical <!--  EDITED_Start responsibilities-->goals<!--  _EDITED_END --> for CHEM <!--  EDIT -->developers and funders<!--  /EDIT --<!-- and criteria for assessing ethical CHEM implementations -->;<!--  MOVED_EDITED_START --> (ii) attributes of CHEMs that can reduce ethical risks; and (iii) criteria against which these attributes can be assessed.  <!--  MOVED_EDITED_END -->  
<!-- (i) some core ethical responsibilities of CHEM project teams;

>>>>

---
title: "MS_Aims"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
<!-- Important rewordings --> 
In this paper, we <!--  MOVED_EDITED_START -->draw on our experience with developing and using CHEMs and consider literature on modelling and software development practice to propose<!--  MOVED_EDITED_END -->:  (i) an initial set of research ethics <!--  EDITED_Start responsibilities-->goals<!--  _EDITED_END --> for CHEM <!--  EDIT -->developers and funders<!--  /EDIT --<!-- and criteria for assessing ethical CHEM implementations -->;<!--  MOVED_EDITED_START --> (ii) attributes of CHEMs that suggest achievement of these goals; and (iii) criteria against which these attributes can be assessed.  <!--  MOVED_EDITED_END -->  
<!-- (i) some core ethical responsibilities of CHEM project teams;

### 2024-03-20 Markdown/Manuscript/Child_RMDs/MS_Discussion.Rmd
ready for preprint

Prior literature has noted that health economic models routinely have limitations in transparency [@Jalali2021; @McManus2019; @Bermejo2017; @Ghabri2019], reuse [@Feenstra2022; @Emerson2019], and being kept up to date [@Sampson2017; @10.3389/fpubh.2022.899874]. These limitations not only pose ethical risks such as those we have described, but are practical barriers to achieving potential exciting enhancements to health economic modelling practice such as improved model transferability [@munthe2019systematic; @barbieri2010international], greater use of reference models [@Afzali2013], living health technology assessments [@thokala2023living] and leveraging the analytic capabilities of machine learning [@padula2022machine].

It is important that the health economics community engages with the ethical dimensions of how health economic models are implemented as software. Recent developments in Large Language Models (LLMs) mean software developers are confronted with novel ethical issues including: increased risk of security flaws; a potential need to add restrictions to open-source licensing; environmentally harmful levels of computing power; and concerns about model explainability and unintended consequences [@ozkaya2023application]. Early work exploring the use of LLMs in CHEM development has already been undertaken [@poirrier2023msr26]. Many of the active ethics debates in the computer science community are therefore likely to become increasing relevant to health economists and public health researchers. Documents such as the World Health Organisation's recent guidance on generative artificial intelligence [@world2024ethics] will be important reference points for how the health economics discipline navigates emerging ethical themes in CHEM development.

Meeting all six TRU criteria we propose is more likely when a CHEM has been implemented as an open-source project<!-- and model transferability -->. It should be noted that the ethical dimensions of open-source software development are broader than the issues we have discussed here, including issues relating to social contracts, contributor autonomy and public goods [@grodzinsky2003ethical]. There are also important barriers to growing the (currently paltry) number of open-source CHEMs. These barriers include health economist concerns about intellectual property, confidential data, the risk of model misuse and the resources required for model maintenance [@Pouwels2022; @Wu2019]. Commercial considerations, for example when health economic models are owned by pharmaceutical companies and consultancies, may also limit the public availability of models and their constituent code and data [@Feenstra2022]. Other barriers relate to <!-- model generalisability and transferability and  -->the software platforms in which CHEMs are authored [@Pouwels2022].

Recent efforts to develop the open-source CHEMs field include an initiative to develop a suite of open source CHEMs addressing different health conditions in the United States [@Jansen2019] and a Canadian platform facilitating access to and validation of multiple CHEMs [@harvard2022developing]. It has also been recommended that training curricula for current and future health economists need to develop some foundational software development skills [@kindilien2021role]. Education efforts to develop such skills could be reinforced with best-practice scientific computing guidance for health economists. Adherence to good practice guidance is an essential requirement for healthcare modelling [@Erdemir2020], but guidelines for implementing open source CHEMs remain scarce, piecemeal and need improving [@Sampson2019]. Health economists also need to be able to access reliable and up to date legal advice about the implications of alternative approaches to software licensing.

@@

Many of these barriers require significant investments by governments and other research funders in enabling infrastructure and human capital. <!-- A core responsibility of research funders is reducing waste in research [@chalmers2014increase]. Investments in modelling projects might generate greater social returns if they supported the project activities necessary to develop CHEMs that are easily understood, widely reused and well maintained.  --> Investment strategies that have previously been recommended for development of the health economic research field include support for harmonised ethical standards for model development [@10.3389/fpubh.2017.00068], methodological innovation to improve model transferability [@craig2018taking], networks of modellers working on common health conditions [@Sampson2019], and  centralized infrastructure such as open-source model repositories [@Pouwels2022] and a standard platform for model implementations [@Ghabri2019].

>>>>

Health economic models routinely have limitations in transparency [@Jalali2021; @McManus2019; @Bermejo2017; @Ghabri2019], reusability [@Feenstra2022; @Emerson2019], and updatability [@Sampson2017; @10.3389/fpubh.2022.899874]. Few current CHEMs would meet all six TRU criteria that we propose, which collectively require that a CHEM is implemented as an open-source project<!-- and model transferability -->. Despite in principle support from many in the health economics field [@Pouwels2022], open-source CHEMs remain rare [@Feenstra2022; @Emerson2019].

Barriers to open-source CHEMs include health economist concerns about intellectual property, confidential data, the risk of model misuse and the resources required for model maintenance [@Pouwels2022; @Wu2019]. Commercial considerations, for example when health economic models are owned by pharmaceutical companies and consultancies, may also limit the public availability of models and their constituent code and data [@Feenstra2022]. Other barriers relate to <!-- model generalisability and transferability and  -->the software platforms in which CHEMs are authored [@Pouwels2022].

<!-- Recent efforts to develop the open-source CHEMs field include an initiative to develop a suite of open source CHEMs addressing different health conditions in the United States [@Jansen2019] and a Canadian platform facilitating access to and validation of multiple CHEMs [@harvard2022developing].  -->
Adherence to good practice guidance is an essential requirement for healthcare modelling [@Erdemir2020], but guidelines for implementing open source CHEMs remain scarce, piecemeal and need improving [@Sampson2019]. Training curricula for current and future health economists that develop foundational software development skills have been recommended [@kindilien2021role]. Health economists also need to be able to access reliable and up to date legal advice about the implications of alternative approaches to software licensing. The ethical dimensions of open-source software development are broader than the issues we have discussed here and include issues relating to social contracts, contributor autonomy and public goods [@grodzinsky2003ethical].

@@

Achieving the ethical goals we propose will require significant investments by governments and other research funders in enabling infrastructure and human capital. <!-- A core responsibility of research funders is reducing waste in research [@chalmers2014increase]. Investments in modelling projects might generate greater social returns if they supported the project activities necessary to develop CHEMs that are easily understood, widely reused and well maintained.  --> Related investment strategies that have previously been recommended for development of the health economic research field include support for harmonised ethical standards for model development [@10.3389/fpubh.2017.00068], methodological innovation to improve model transferability [@craig2018taking], networks of modellers working on common health conditions [@Sampson2019], and  centralized infrastructure such as open-source model repositories [@Pouwels2022] and a standard platform for model implementations [@Ghabri2019].

### 2024-03-20 Markdown/Manuscript/Child_RMDs/MS_Imp_Guidelines.Rmd
ready for preprint

Boden and McKendrick [@10.3389/fpubh.2017.00068] propose a framework for ethical public health modelling based on the criteria of *independence* (concerning how modeller subjectivity shapes model design), *beneficence / non-maleficence* (concerning model quality and utility), *transparency* (concerning the need for policymakers to reliably evaluate model strengths and weaknesses) and *justice* (concerning the social obligations of modellers to consider and communicate ethical issues about model use). <!-- However, t -->This framework is not specific to a model's implementation as software, but <!-- are designed to  -->apply to a modelling project as a whole, including a model's conceptual and mathematical representations. Boden and McKendrick proposed 13 questions to evaluate ethical risk across the four criteria, but of these, only two pertain to computational models: "is the model code open-source or available on request" (for transparency) and "has the model been verified, i.e., does it do what the modeller wants it to do?" (for beneficence / non-maleficence).

To identify ethical issues that are specific to CHEMs, we <!-- first  -->considered <!-- Edit -->how Resnick's 18 principles and Boden and McKendrick's four criteria might apply to <!-- /Edit -->the distinct phases in a software development project. Software development lifecycle (SDLC) models take different approaches to the inclusion, naming, definition, sequencing and iteration of project phases [@ruparelia2010software]. However, SDLC models typically have components that can also principally map to the concepts of *planning* (e.g., identification of system / user requirements and design specifications), *implementation* (e.g., development and testing) and *release* (e.g., deployment, system integration, maintenance and support). For simplicity, we propose one overarching ethical <!-- responsibility  for CHEM project team-->goal for each phase. These goals<!-- responsibilities  --> are: 
<!-- establish  -->
(i) *socially acceptable user requirements and design specifications* (during CHEM planning); 
<!-- ensure  -->
(ii) *fitness for purpose* (of CHEM implementations); and 
<!-- support  -->
(iii) *socially beneficial use* (post CHEM release).

@@

Misalignment between the values of computational model developers and those of the population groups affected by decisions based on their models presents significant ethical risks [@thompson2022escape; @duckett2022journey].  The value judgments of health economic modellers are rarely adequately specified [@duckett2022journey]. These value judgments influence the assumptions, selection of model features and standards for evidence that shape health economic model projects [@HARVARD2020112975]. For example, the process of determining CHEM *user requirements and design specifications* will involve making judgments about the competing claims of open-source and commercial development platforms (Table \@ref(tab:proscons)). That decision will in turn determine how straightforward it is for a CHEM to be subsequently amended to reflect alternative value judgments relating to model conceptualisation and mathematical formalisation (Table \@ref(tab:timelygls)).
<!-- Moved -->
<!--  , omissions that may lead to socially unacceptable policy recommendations -->

@@

<!-- Moved -->
<!-- Edit --><!-- All three  -->Our proposed ethical goals are most likely to be met when there is both good<!-- have implications for both  --> professional practice by CHEM developers and <!-- Addition -->adequate <!-- Addition -->resourcing by project funders. <!-- /Edit --> The choices CHEM developers make often need to balance ethical considerations with project resource constraints (as illustrated in Table \@ref(tab:proscons)). Modelling projects implemented by diligent and conscientious CHEM developers may still be exposed to ethical risks if there are inadequate resources to meaningfully engage important stakeholder groups, prepare code and data artefacts for independent assessment and provide ongoing model maintenance and user support. Currently, it can take “an extraordinary amount of idealism” to commit to authoring and maintaining research software [@anzt2020environment]. Existing incentive structures for research health economists generally do not promote facilitating peers to reuse their work and funding is rarely adequate to maintain and routinely update a model. Economists may therefore make pragmatic judgments that proprietary licensing approaches provide the most feasible pathway to mobilising the resources to support model maintenance. For these reasons, we believe responsibility for achieving our proposed ethical goals is shared between CHEM developers and funders. Reflecting the shared nature of this responsibility, we suggest that in most cases where a modelling project fails to meet our ethical goals, it will be more appropriate to describe these shortcomings as "ethical risks" (the language used by Boden and McKendrick) than to use accusatory terms like "unethical practice."<!-- these responsibilities should apply jointly to health economic model developers and funders. -->

>>>>

Boden and McKendrick [@10.3389/fpubh.2017.00068] propose a framework for ethical public health modelling based on the criteria of *independence* (concerning how modeller subjectivity shapes model design), *beneficence / non-maleficence* (concerning model quality and utility), *transparency* (concerning the need for policymakers to reliably evaluate model strengths and weaknesses) and *justice* (concerning the social obligations of modellers to consider and communicate ethical issues about model use). <!-- However, t -->This framework is not specific to a model's implementation as software, but <!-- are designed to  -->applies to a modelling project as a whole, including a model's conceptual and mathematical representations. Boden and McKendrick proposed 13 questions to evaluate ethical risk across the four criteria, but of these, only two pertain to computational models: "is the model code open-source or available on request" (for transparency) and "has the model been verified, i.e., does it do what the modeller wants it to do?" (for beneficence / non-maleficence).

To identify ethical issues that are specific to CHEMs, we <!-- first  -->considered <!-- Edit -->how Resnick's 18 principles and Boden and McKendrick's four criteria might apply to <!-- /Edit -->the distinct phases in a software development project. Software development lifecycle (SDLC) models take different approaches to the inclusion, naming, definition, sequencing and iteration of project phases [@ruparelia2010software]. However, SDLC models typically have components that can principally map to the concepts of *planning* (e.g., identification of system / user requirements and design specifications), *implementation* (e.g., development and testing) and *release* (e.g., deployment, system integration, maintenance and support). For simplicity, we propose one overarching ethical <!-- responsibility  for CHEM project team-->goal for each phase. These goals<!-- responsibilities  --> are: 
<!-- establish  -->
(i) *socially acceptable user requirements and design specifications* (during CHEM planning); 
<!-- ensure  -->
(ii) *fitness for purpose* (of CHEM implementations); and 
<!-- support  -->
(iii) *socially beneficial use* (post CHEM release).

@@

Misalignment between the values of computational model developers and those of the population groups affected by decisions based on their models presents significant ethical risks [@thompson2022escape; @duckett2022journey].  The value judgments of health economic modellers are rarely adequately specified [@duckett2022journey]. These value judgments influence the assumptions, selection of model features and standards for evidence that shape health economic model projects [@HARVARD2020112975]. For example, the process of determining CHEM *user requirements and design specifications* will involve making judgments about the competing claims of proprietary and open-source development platforms (Table \@ref(tab:proscons)). An advantage of commercial <!-- modelling  -->software is that users often require limited or no programming skills to develop and apply CHEMs.  <!-- Addition --><!-- /Addition --> Open-source software programming languages, on the other hand, may facilitate the development of CHEMs that are more transparent, reusable and updatable [@incerti2019r; @Pouwels2022]. The choice of software development platform can influence how easily a CHEM can be amended by others to reflect alternative value judgments relating to model conceptualisation and mathematical formalisation (Table \@ref(tab:timelygls)).
<!-- Moved -->
<!--  , omissions that may lead to socially unacceptable policy recommendations -->

@@

<!-- Moved -->
<!-- Edit --><!-- All three  -->Our proposed ethical goals are most likely to be met when there is both good<!-- have implications for both  --> professional practice by CHEM developers and <!-- Addition -->adequate <!-- Addition -->resourcing by project funders. <!-- /Edit --> The choices CHEM developers make often need to balance ethical considerations with project resource constraints (Table \@ref(tab:proscons)). Modelling projects implemented by diligent and conscientious CHEM developers may still be exposed to ethical risks if there are inadequate resources to meaningfully engage important stakeholder groups, prepare code and data artefacts for independent assessment and provide ongoing model maintenance and user support. Currently, it can take “an extraordinary amount of idealism” to commit to authoring and maintaining research software [@anzt2020environment]. Existing incentive structures for research health economists generally do not promote facilitating peers to reuse their work and funding is rarely adequate to routinely update a model. Economists may therefore make pragmatic judgments that proprietary licensing approaches provide the most feasible pathway to mobilising the resources to support model maintenance. For these reasons, we believe responsibility for achieving our proposed ethical goals is shared between CHEM developers and funders. Reflecting the shared nature of this responsibility, we suggest that in most cases where a modelling project fails to meet our ethical goals, it will be more appropriate to describe these shortcomings as "ethical risks" (the language used by Boden and McKendrick) than to use accusatory terms like "unethical practice."<!-- these responsibilities should apply jointly to health economic model developers and funders. -->

### 2024-03-20 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
ready for preprint

The set of software files implementing a health economic model's structure, data and algorithms can be called a computational health economic model (CHEM). CHEMs can be developed using commercial modelling <!-- Addition -->or spreadsheet<!-- /Addition --> software (for example, Microsoft Excel [@msexcel]) or open-source programming languages such as Python [@python2009] or R [@RCORE2022]. An advantage of commercial <!-- modelling  -->software is that users often require limited or no programming skills to develop and apply CHEMs.  <!-- Addition -->It has been argued that<!-- /Addition --> open-source software programming languages may facilitate the development of CHEMs that are more transparent, reusable and updatable [@incerti2019r; @Pouwels2022]. <!-- Moved -->However, open-source CHEMs remain rare [@Feenstra2022; @Emerson2019]<!-- Moved -->.

Health economists must balance ethical considerations and resource constraints when deciding what type of software development platform to use for authoring CHEMs. <!-- ( see--><!-- Addition --> In Table \@ref(tab:proscons), we summarise some of the trade-offs we consider when undertaking modelling projects<!-- ) -->. <!-- Edit Move -->Determining how to make these trade-offs can be difficult. The computational modelling landscape is changing fast [@calder2018computational] and the ethical <!-- responsibilities  --><!-- Edit -->dimensions of computational public health modelling <!-- /Edit -->remain poorly understood [@10.3389/fpubh.2017.00068]. <!-- /Edit Move-->

>>>>

The set of software tools and files implementing a health economic model's structure, data and algorithms can be called a computational health economic model (CHEM). CHEMs can be developed using commercial modelling <!-- Addition -->or spreadsheet<!-- /Addition --> software (for example, Microsoft Excel [@msexcel]) or open-source programming languages such as Python [@python2009] or R [@RCORE2022]. Training and professional practice guidance for health economists typically provide only superficially coverage of software development topics. However, these capabilities will be critical for achieving proposed methodological innovations at the frontier of the health economics field. Opportunities for field development that closely overlap with CHEM implementation approaches include living health technology assessments [@thokala2023living], machine learning powered health economic analyses [@padula2022machine], open source health-economic models [@Pouwels2022] and utilising Large Language Models (LLMs) to program health economic models [@poirrier2023msr26]. <!-- Potential enhancements to health economic modelling practice include improved model transferability [@munthe2019systematic; @barbieri2010international], greater use of reference models [@Afzali2013], living health technology assessments [@thokala2023living] and leveraging the analytic power of machine learning [@padula2022machine]. -->

There is a significant ethical dimension to software development. For example, recent developments in LLMs mean software developers may confront novel ethical issues including: new security risks; more restrictive open-source licensing; environmentally harmful levels of computing power; and concerns about model explainability and unintended consequences [@ozkaya2023application]. The health economics community should engage more directly with these ethical debates.  The computational modelling landscape is changing fast [@calder2018computational] and the ethical <!-- responsibilities  --><!-- Edit -->dimensions of computational public health modelling <!-- /Edit -->remain poorly understood [@10.3389/fpubh.2017.00068]. Documents such as the World Health Organisation's recent guidance on generative artificial intelligence [@world2024ethics] can be important reference points in helping the health economics discipline navigate relevant ethical topics in software development.

  <!-- Moved -->
 <!-- However, open-source CHEMs remain rare [@Feenstra2022; @Emerson2019], despite in principle support from many in the health economics field [@Pouwels2022]. -->
 <!-- Moved -->

<!-- Health economists must balance ethical considerations and resource constraints when deciding what type of software development platform to use for authoring CHEMs. <!-- ( see--><!-- Addition --> 
<!--In Table \@ref(tab:proscons), we summarise some of the trade-offs we consider when undertaking modelling projects<!-- ) --> <!-- Edit Move --><!--Determining how to make these trade-offs can be difficult.  -->

 <!-- /Edit Move-->

### 2024-03-20 Markdown/Manuscript/Child_RMDs/MS_Tables_2.Rmd
ready for preprint

---
title: "Predr_Cors"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
```{r timelygls, tab.cap='Transparent, reusable and updatable (TRU) computational health economic models (CHEMs) are suggestive of ethical modelling practice', tab.id = 'timelygls', echo = F, results = 'asis', warning=FALSE}
library(kableExtra)  
df <- data.frame(` ` = c("Transparent", "Reusable", "Updatable"),
                 `Social acceptability` = c("Modeller judgments, model features and verification checks can be reviewed by third parties.",
                                                                        "Models can be modified to reflect alternative value judgments.",
                                                                        "Models can be modified to reflect alternative value judgments."),
                 `Fitness for purpose` = c("Modeller judgments, model features and verification checks can be reviewed by third parties.","Third party use increases likelihood of errors being identified.","Models can be validly used for longer."),
                 `Beneficial use` = c("Decision-makers can understand the strengths and weakness of models before applying them.","Can inform more decision problems, in more contexts with less duplicative effort.",""#"Models can be validly used for longer."
                                      ),
                 check.names = F)
 library("flextable")
  library("ftExtra")
if(params$X@outp_formats_chr[1] == "Word") {

  df %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    # library(huxtable)
    # library(dplyr)
   # flextable::set_flextable_defaults() #fonts_ignore=TRUE
    flextable::flextable(df) %>%
      # flextable::border_inner(border = officer::fp_border(style = "none")) %>%
      # flextable::border_outer() %>%
      flextable::align(align = "center", part = "header") 
    # %>%
    #   flextable::merge_at(i = 3, j = 3:4)
      # flextable::merge_h() %>%
      # flextable::merge_v() %>%
      # flextable::valign(j=2, valign = "top") %>%
      # flextable::align(j=2:4, align = "center") %>%
      # flextable::bold(j=1, part = "all") %>%
      # flextable::bold(part = "header") %>%
      # flextable::width(j = 1, width = 1.2) %>%
      # flextable::width(j = 2:4, width = 1.6)
    # %>%
    #   
      #flextable::merge_at(i=2,j=1:2)
    # huxtable::as_hux(df) %>% 
    #   #insert_row("MPG-DISP", "", "", "HP-WT", fill = "") %>% 
    #   #huxtable::merge_cells(2:3,1) %>% 
    #  # huxtable::set_header_rows(1, TRUE) %>% 
    #   huxtable::set_col_width(c(.1,.1,.1,.7)) %>%
    #   huxtable::set_wrap(1, value=T)
      #huxtable::style_headers(bold = TRUE)
    # df %>%
    #   kableExtra::kbl(booktabs = T,
    #                   escape = F,
    #                   caption = knitr::opts_current$get("tab.cap")) %>%
    #   kableExtra::kable_paper(full_width = F) %>%
    #   kableExtra::column_spec(1, bold=T) %>%
    #   kableExtra::add_header_above(c(" ", "Project team responsibility" = 3), bold = T, italic = T) %>%
    #   kableExtra::column_spec(2, width = "13em") %>%
    #   kableExtra::column_spec(3, width = "13em") %>%
    #   kableExtra::column_spec(4, width = "13em") %>%
    #   kableExtra::collapse_rows(columns = 2, latex_hline = "none")

    }
```

>>>>

---
title: "Predr_Cors"
author: "Author"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r echo=FALSE}
df <- data.frame(` ` = c("Transparent", "Reusable", "Updatable"),
                 `Social acceptability` = c("Modeller judgments, model features and verification checks can be reviewed by third parties.",
                                                                        "Models can be modified to reflect alternative value judgments.",
                                                                        ""#"Models can be modified to reflect alternative value judgments."
                                            ),
                 `Fitness for purpose` = c("",#"Modeller judgments, model features and verification checks can be reviewed by third parties.",
                                           "Third party use increases likelihood of errors being identified.",
                                           "Models can be validly used for longer."),
                 `Beneficial use` = c("Decision-makers can understand the strengths and weakness of models before applying them.",
                                      "Can inform more decision problems, in more contexts with less duplicative effort.",
                                      ""#"Models can be validly used for longer."
                                      ),
                 check.names = F)
```
```{r echo=FALSE}
if(params$X@outp_formats_chr[1] == "Word"){
  conditional_1L_chr <- "../Child_RMDs/MS_Tables_2_WORD.Rmd"
}else{
    conditional_1L_chr <- "../Child_RMDs/MS_Tables_2_LATEX.Rmd"
}
```
```{r child = conditional_1L_chr, echo=FALSE}
```

### 2024-03-20 Markdown/Manuscript/Parent_PDF/Springer.Rmd
v4
ready for preprint

title: Ethical considerations for computational implementation of health economic models
titlerunning: Ethical computational modelling

@@

affiliations:
  - number: 1
    corresponding: TRUE
    info:
      orgdiv: School of Public Health and Preventive Medicine
      orgname: Monash University
    address:
        #street: "553 St Kilda Road"
        city: Melbourne
        #postcode: 3004
        #state: Victoria
        country: Australia
  - number: 2
    corresponding: FALSE
    info:
      orgname: Orygen
    address:
     #   street: "35 Poplar Road"
        city: Parkville
        # postcode: 3052
        #state: Victoria
        country: Australia
  - number: 3
    corresponding: FALSE
    info:
#      orgdiv: XXXXXXXXXXX
      orgname: Flinders University   
    address:
    #     # street: "182 Victoria Square"     
        city: Adelaide
    #     postcode: 5000
    #  #   state: South Australia
        country: Australia 
  - number: 4
    corresponding: FALSE
    info:
      orgdiv: Centre for Youth Mental Health
      orgname: The University of Melbourne
    address:
     #   street: "35 Poplar Road"
        city: Parkville
        # postcode: 3052
    #   state: Victoria
        country: Australia
#   - number: 4
#     corresponding: FALSE
#     info:
# #      orgdiv: XXXXXXXXXXX
#       orgname: Heart Foundation      
#     address:
#      #   street: "2/850 Collins St"
#         city: Docklands
#         # postcode: 3008
#      #   state: Victoria
#         country: Australia
#   - number: 5
#     corresponding: FALSE
#     info:
# #      orgdiv: XXXXXXXXXXX
#       orgname: headspace National Youth Mental Health Foundation   
#     address:
#     #  #   street: "485 La Trobe St"
#         city: Melbourne
#     #     postcode: 3000
#     #  #   state: Victoria
#         country: Australia     
#   - number: 6
#     corresponding: FALSE
#     info:
#       orgdiv: Institute for Health and Sport
#       orgname: Victoria University 
#     address:
#     #  #   street: "70/104 Ballarat Rd"
#         city: Footscray
#     #     postcode: 3011
#     #  #   state: Victoria
#         country: Australia 
#       # 
  - number: 5
    corresponding: FALSE
    info:
#      orgdiv: XXXXXXXXXXX
      orgname: University of Canberra  
    address:
    #     # street: "182 Victoria Square"     
        city: Canberra
    #     postcode: 5000
    #  #   state: South Australia
        country: Australia 
keywords:  
  - computational models
  - ethics of modelling
  - health economics
  - open-source models

@@

  \usepackage{float}
  \let\origfigure\figure
  \let\endorigfigure\endfigure
  \renewenvironment{figure}[1][2] {
      \expandafter\origfigure\expandafter[H]
  } {
      \endorigfigure
  }
  \usepackage{lscape}
  \usepackage{pdflscape}
  \newcommand{\blandscape}{\begin{landscape}}
  \newcommand{\elandscape}{\end{landscape}}
  \usepackage[none]{hyphenat}
  \usepackage[export]{adjustbox}
  \usepackage{multirow}
  \usepackage{longtable}
  \usepackage{array}
  \usepackage{wrapfig}
  \usepackage{colortbl}
  \usepackage{booktabs} 
  \usepackage{colortbl}
  \usepackage{pdflscape}
  \usepackage{tabu}
  \usepackage{threeparttable}
  \usepackage[normalem]{ulem}
  \usepackage{threeparttablex}
  \usepackage{makecell}

>>>>

title: "Ethical considerations when planning, implementing and releasing health economic model software: a new proposal"
titlerunning: Ethical modelling software

@@

affiliations:
  - number: 1
    corresponding: TRUE
    info:
      orgdiv: School of Public Health and Preventive Medicine
      orgname: Monash University
    address:
        #street: "553 St Kilda Road"
        city: Melbourne
        #postcode: 3004
        #state: Victoria
        country: Australia
  - number: 2
    corresponding: FALSE
    info:
      orgname: Orygen
    address:
     #   street: "35 Poplar Road"
        city: Parkville
        # postcode: 3052
        #state: Victoria
        country: Australia
  - number: 3
    corresponding: FALSE
    info:
      orgdiv: Flinders Health and Medical Research Institute
      orgname: Flinders University   
    address:
    #     # street: "182 Victoria Square"     
        city: Adelaide
    #     postcode: 5000
    #  #   state: South Australia
        country: Australia 
  - number: 4
    corresponding: FALSE
    info:
      orgdiv: Centre for Youth Mental Health
      orgname: The University of Melbourne
    address:
     #   street: "35 Poplar Road"
        city: Parkville
        # postcode: 3052
    #   state: Victoria
        country: Australia
#   - number: 4
#     corresponding: FALSE
#     info:
# #      orgdiv: XXXXXXXXXXX
#       orgname: Heart Foundation      
#     address:
#      #   street: "2/850 Collins St"
#         city: Docklands
#         # postcode: 3008
#      #   state: Victoria
#         country: Australia
#   - number: 5
#     corresponding: FALSE
#     info:
# #      orgdiv: XXXXXXXXXXX
#       orgname: headspace National Youth Mental Health Foundation   
#     address:
#     #  #   street: "485 La Trobe St"
#         city: Melbourne
#     #     postcode: 3000
#     #  #   state: Victoria
#         country: Australia     
#   - number: 6
#     corresponding: FALSE
#     info:
#       orgdiv: Institute for Health and Sport
#       orgname: Victoria University 
#     address:
#     #  #   street: "70/104 Ballarat Rd"
#         city: Footscray
#     #     postcode: 3011
#     #  #   state: Victoria
#         country: Australia 
#       # 
  - number: 5
    corresponding: FALSE
    info:
      orgdiv: Health Research Institute
      orgname: University of Canberra  
    address:
    #     # street: "182 Victoria Square"     
        city: Canberra
    #     postcode: 5000
    #  #   state: South Australia
        country: Australia 
keywords:  
  - computational models
  - ethics of modelling
  - health economics
  - open-source models

@@

  \usepackage{float}
  \let\origfigure\figure
  \let\endorigfigure\endfigure
  \renewenvironment{figure}[1][2] {
      \expandafter\origfigure\expandafter[H]
  } {
      \endorigfigure
  }
  \usepackage{lscape}
  \usepackage{pdflscape}
  \newcommand{\blandscape}{\begin{landscape}}
  \newcommand{\elandscape}{\end{landscape}}
  \usepackage[none]{hyphenat}
  \usepackage[export]{adjustbox}
  \usepackage{multirow}
  \usepackage{longtable}
  \usepackage{array}
  \usepackage{wrapfig}
  \usepackage{colortbl}
  \usepackage{booktabs} 
  \usepackage{colortbl}
  \usepackage{pdflscape}
  \usepackage{tabu}
  \usepackage{threeparttable}
  \usepackage[normalem]{ulem}
  \usepackage{threeparttablex}
  \usepackage{makecell}
  \usepackage{array}
  \newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
  \newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
  \newcolumntype{N}{@{}m{0pt}@{}}
  \usepackage{caption}
  \usepackage{graphicx}
  \usepackage{siunitx}
  \usepackage{hhline}
  \usepackage{calc}
  \usepackage{tabularx}
  \renewcommand\tabularxcolumn[1]{m{#1}}% for vertical centering text in X column

  \usepackage{threeparttable}
  \usepackage{wrapfig}
  \usepackage{adjustbox}
  \usepackage{hyperref}
